manuscript_removed_text

### 2019-08-30 manuscript/manuscript_removed_text.Rmd
Update manuscript_removed_text.Rmd

>>>>

Globally, forests play an important role in the carbon cycle and are a major component of global carbon dioxide budgets [@luyssaert_old-growth_2008]. They show higher levels of productivity than non-forest terrestrial ecosystems [@del_grosso_global_2008], and as a result achieve significant carbon sequestration and storage. Estimating the total role of forests in the carbon cycle is challenging, but studies indicate that old growth forests alone sequester up to 1.4 GtCyr^-1^ [@malhi_carbon_1999], while the total sequestration of carbon by established forests globally could be up to 2.4 GtCyr^-1^, with the largest sinks being in old-growth tropical forests [@pan_large_2011]. As atmospheric carbon dioxide levels continue to rise, with consequences for global climate, there is increasing recognition that proper protection and management of forest resources will have an important role to play in mitigating climate change. Understanding the patterns of forest productivity on a global scale, and the drivers behind them, is therefore a priority in forest research.

There are two major questions to understand: firstly, how forest productivity varies globally - and specifically which areas show the greatest peaks in productivity -; and secondly, which climate variables drive this variation. On a global scale, the productivity of forests varies with latitude, with a general trend of increasing productivity towards the tropics [@beer_terrestrial_2010; @jung_global_2011]; however the exact nature of this pattern, and how it varies by component of productivity, is poorly understood. This latitudinal gradient is most likely to be explained by climatic gradients in temperature, precipitation, length of growing season, and combinations of the above. Productivity is influenced by a range of climatic drivers, including mean annual temperature (MAT) and mean annual precipitation (MAP) [@del_grosso_global_2008], but doesn't necessarily respond linearly to these drivers. Disentangling the shape of productivity responses to climate drivers will enable better predictions of future responses under climate change.

*The components of forest productivity.* During photosynthesis, plants capture carbon dioxide from the atmosphere. The gross primary productivity (GPP) of an ecosystem is the gross uptake, via photosynthesis, of carbon dioxide by plants in that ecosystem. Only a fraction of the carbon captured is assimilated into plant tissue; the rest is used in autotrophic respiration (R~auto~). The component of GPP that is stored as plant material is the net primary productivity (NPP) of an ecosystem. Net primary productivity can therefore be expressed as:

$$ NPP = GPP + R~auto~ $$

Currently, GPP cannot be measured directly by observing total ecosystem photosynthesis. Instead field estimates of GPP have to be derived based on modelling and extrapolation of eddy-covariance studies and measurements of net ecosystem exchange (NEE) [@clark_reviews_2017].

In contrast, NPP can be calculated through direct field observations. In order to achieve greater accuracy in estimating NPP, NPP is often broken down into its component parts, with aboveground NPP (ANPP) and belowground NPP (BNPP) considered separately to each other. The components included when estimating ANPP and BNPP often vary between studies, depending on the intensity of fieldwork effort. At its most basic level, ANPP can be expressed as:

$$ ANPP = NPP~stem~ + NPP~branch~ + NPP~canopy~ $$
where NPP~stem~ is the annual woody increment of all stems above a specified diameter at breast height (DBH), NPP~branch~ is annual woody branch turnover, and NPP~canopy~ is annual foliage production, including leaves, twigs, and reproductive structures. ANPP may also include NPP~VOC~, the annual emission of volatile organic compounds, and NPP~herbivory~, the annual consumption of plant matter by herbivores, but these components are often excluded from field observations as they are much harder to quantify. Other components of aboveground productivity that remain largely unquantified include epiphytes, hemiepiphytes, and understory plants [@clark_reviews_2017]. The majority of current ANPP estimates are based on the assumption that the contribution of these components to overall NPP is insignificant.

There are two major subcomponents of BNPP, which can be expressed as:

$$ BNPP = NPP~coarse root~ + NPP~fine root~ $$
where NPP~coarse root~ is the annual production of coarse roots (typically roots >2mm diameter), and NPP~fine root~ is the annual production of fine roots (typically roots <2mm diameter) [@aragao_above-_2009]. Calculations of BNPP may also include NPP~exudation~, a measure of annual carbon losses through root exudation, and NPP~symbionts~, the annual carbon allocation to mycorrhizae and legumes, but, as before, this is challenging to quantify and is often excluded from field observations.

BNPP is a poorly understood component of total ecosystem productivity, primarily because of the challenges in gaining accurate field measurements. Coarse root productivity is often estimated via extrapolation of NPP~stem~ estimates using allometries that may not have been empirically verified [@clark_net_2001]. NPP~fine root~ is easier to quantify through soil cores and minirhizotrons, however, sampling tends to be limited to the surface soils, with very few studies sampling to depths below 3 metres [cite]. As a result, it is possible that BNPP is currently significantly underestimated, despite being a hugely significant component of total ecosystem productivity [@pan_large_2011].

*Which factors influence primary productivity?* Primary productivity is influenced by many factors, which often act across a range of scales, and may show interactive effects with each other. On a local scale, stand age [@litton_carbon_2007; @gillman_latitude_2015], management [@simova_enigma_2017]; nutrient availability [@aragao_above-_2009]; and altitude [@girardin_net_2010; @malhi_variation_2017] all impact forest productivity. On a global scale, changes in primary productivity are influenced by climatic variables and abiotic gradients, such as the length of growing season [@michaletz_convergence_2014]. There is some debate over the precise relationship between these drivers and productivity; while MAT and MAP have been argued to be significant predictors of productivity [@chu_does_2016], other studies have found that the correlation between productivity and MAT is a factor of the relationship between productivity and growing season length [@kerkhoff_plant_2005; @malhi_productivity_2012; @michaletz_convergence_2014; @michaletz_drivers_2018]. Improving understanding of how these factors interact to control global patterns in primary productivity is essential to understanding the global carbon cycle.

Current research into how primary productivity varies with latitude is inconclusive, and - though it has primarily focussed on patterns of GPP, NPP, and ANPP - indicates that different components of productivity may show different relationships with latitude. Gross primary productivity is generally thought to be highest in the tropics. Modelling of global terrestrial ecosystem GPP though upscaling and calibration of eddy flux measurements indicates a peak in GPP in the tropics, with the highest levels in tropical forests [@beer_terrestrial_2010; @jung_global_2011]. This is corroborated by analysis of site-level GPP measurements, which show a strong positive correlation between GPP and MAT and MAP [@luyssaert_co2_2007], with the highest GPP values reported in tropical forests. The influence of latitude on global patterns of NPP is less clear than that of GPP. Simova and Storch (2017) found that, as with GPP, NPP decreases with latitude, peaking in the tropics. However, other studies have found the highest values of NPP in temperate forests [@luyssaert_co2_2007; @huston_global_2009]. Because of the challenges in accurately obtaining unbiased measures of belowground productivity, many studies focus on ANPP in preference to measures of NPP. Studies on global patterns of ANPP are equally inconclusive: Gillman et al. (2015) found a weak negative relationship between ANPP and latitude, with the relationship becoming stronger in older forest stands. These findings were echoed in other studies, which have found weak or no relationships between ANPP and latitude [@huston_global_2009].

Furthermore, there is evidence that different components of productivity show individual responses to climate drivers. For example, increases in GPP have been reported to saturate above 25$^\circ$C MAT [@larjavaara_temperature_2012], while increases in NPP are recorded to saturate above 10$^\circ$C MAT [@luyssaert_co2_2007]. Similarly, allocation to different components of ANPP varies with climate. Within the tropics, allocation to canopy NPP appears fairly consistent, with significantly greater variation in allocation to woody and belowground NPP [@litton_carbon_2007; @malhi_allocation_2011]. Allocation to these structural biomass components has been shown to increase with water availability [@litton_carbon_2007; @bloom_decadal_2016], and is highest in the wet tropics, indicating that control of woody productivity by MAP may be more significant than control of foliar productivity. However, these studies are regional, meaning that our understanding of variation in allocation and its relationship to climate on a global scale remains limited. In addition, allocation is also influenced by stand age [@de_lucia_forest_2007], nutrient availability [@litton_carbon_2007] and forest structure [@taylor_greater_2019], which can make it challenging to disentangle the effects of climate.

Data that control for stand age and standardize methodologies are required to resolve this question. Here, we use a comprehensive global database of forest carbon fluxes to explore global patterns in productivity. We explore five broad questions, described in greater detail in table 1:  
  1. How do carbon fluxes vary with latitude?  
  2. How do carbon fluxes relate to MAT and MAP?  
  3. How do carbon fluxes relate to other annual climate variables?  
  4. What is the role of seasonality in explaining variation in carbon fluxes?  
  5. Per month of the growing season, how does productivity vary with climate?
